---
title: "Plotting"
output:
editor_options:
  chunk_output_type: console
---

In this practical, you will visualize the `sdg_hackathon_data.zip` data set using the `ggplot2` package.

**To complete the practical, you will have to replace the `XX` in the code with the correct object or functions and then run the code in the console.** 

# 0 - Preliminaries

1. If you haven't done so already, load the `tidyverse` and `lubridate` packages and read in the `sdg_hackathon_data.zip` data set. 

```{r}
library(tidyverse)
library(lubridate)
project_sdgs <- read_csv('1_Data/sdg_hackathon_data.zip')
```

# 1 - Wrangling

1. Before you begin with plotting, use the code below to determine the relative proportion of SDG 1 and SDG 6 for the years `2009` to `2021` and the `"elsevier"` system.  

```{r}
# determine relative proportion of sdg 1 and 6 
sdg_props <- project_sdgs %>% 
  mutate(year = year(start_date)) %>% 
  group_by(system, sdg, year) %>% 
  summarize(n = n()) %>% 
  filter(year > 2008 & year < 2022,
         system == "elsevier",
         sdg %in% c("SDG-01", "SDG-06")) %>% 
  group_by(system, year) %>% 
  mutate(prop = n / sum(n))
```

# 2 - Plotting relative SDG proportions

1. Start a ggplot using `ggplot()` by piping the newly created `sdg_props` tibble into the function's `data` argument. 

```{r}
# start plot
sdg_props %>% 
  ggplot()
```

2. Running the previous chunk should have created a blank canvas in your bottom right RStudio panel. Now set up the structure of the plot using the `mapping` argument and the `aes()` helper function. Place `year` on the x-axis and `prop` on the y-axis. 

```{r}
# setup aesthetic mapping
sdg_props %>% 
  ggplot(mapping = aes(x = XX, 
                       y = XX))
```

3. The updated plot should now contain axis labels, tick marks, labels, and grid lines. It's time to add some geometric shapes to actually show the data. Add `geom_point()`. 

```{r}
# add a geom
sdg_props %>% 
  ggplot(mapping = aes(x = XX, 
                       y = XX)) +
  XX
```

4. The figure should show two sets of points, one set increasing with year and one set decreasing with year. Try adding an additional `geom_line()` to emphasize the trends.  

```{r}
# add a second geom 
sdg_props %>% 
  ggplot(mapping = aes(x = XX, 
                       y = XX)) +
  XX + XX
```

5. That didn't work as intended. The reason is that `ggplot` currently doesn't now that these points belong to two different groups. One way to communicate this and to introduce colors into the plot at the same time is to make use of the `color` aesthetic in the mapping by assigning `sdg` to it. 

```{r}
# using the color aesthetic
sdg_props %>% 
  ggplot(mapping = aes(x = XX, 
                       y = XX,
                       color = XX)) +
  XX + XX
```

6. That should have done the trick. Do you think a `geom_smooth()` would make the plot better?

# 3 - Facetting

1. Facetting allows you to automatically create the same plot for the values of a (discrete) grouping variable, such as for instance the SDG labeling `system`. Let's try it out. First, use the code below to determine the relative proportion of SDGs 1 and 6 for each system.   

```{r}
# determine relative proportion of sdg 1 and 6 
sdg_props_systems <- project_sdgs %>% 
  mutate(year = year(start_date)) %>% 
  group_by(system, sdg, year) %>% 
  summarize(n = n()) %>% 
  
  # no filtering of systems
  filter(year > 2008 & year < 2022,
         sdg %in% c("SDG-01", "SDG-06")) %>% 
  
  # group by year and system
  group_by(year, system) %>%
  mutate(prop = n / sum(n))
```

2. Now use the newly created `sdg_props_systems` tibble to create the same plot as before, but now facetted according to `system`.

```{r}
# plot facets
sdg_props_systems %>% 
  ggplot(mapping = aes(x = XX, 
                       y = XX,
                       color = XX)) +
  XX + XX +
  facet_wrap(~ XX)
```









