---
title: "R Notebook"
output:
editor_options:
  chunk_output_type: console
---

In this practical, you will visualize the `sdg_hackathon_data.zip` data set using the `ggplot2` package and explore the packages basic styling functionality.

**To complete the practical, you will have to replace the `XX` in the code with the correct object or functions and then run the code in the console.** 

# 0 - Preliminaries

1. If you haven't done so already, load the `tidyverse` and `lubridate` packages and read in the `sdg_hackathon_data.zip` data set. Additionally, load the `ggthemes` package. 

```{r}
library(tidyverse)
library(lubridate)
library(ggthemes)
project_sdgs <- read_csv('1_Data/sdg_hackathon_data.zip')
```

# 1 - Basic plot

1. To start with, use the wrangling code below to determine the relative proportion of SDG 1 and 6 for years 2008 to 2021 and the `"elsevier"` system.   

```{r}
# determine relative proportion of sdg 1 and 6 
sdg_props <- project_sdgs %>% 
  mutate(year = year(start_date)) %>% 
  group_by(system, sdg, year) %>% 
  summarize(n = n()) %>% 
  filter(year > 2008 & year < 2022,
         system == "elsevier",
         sdg %in% c("SDG-01", "SDG-06")) %>% 
  group_by(system, year) %>% 
  mutate(prop = n / sum(n))
```

2. Recreate the plot from the previous session showing the relative proportions of the two SDGs 1 and 6 in different colors using both points (`geom_point()`) and lines (`geom_line()`). 

```{r}
# recreate plot
sdg_props %>% 
  ggplot(mapping = aes(x = XX, 
                       y = XX,
                       color = XX)) +
  XX + XX
```

3. Now save the plot in an object called `sdg_plot` to simplify the next steps. 

```{r}
# recreate plot
sdg_plot <- sdg_props %>% 
  ggplot(mapping = aes(x = XX, 
                       y = XX,
                       color = XX)) +
  XX + XX
```


# 2 - Themes

1. Add `theme_minimal()` to your `sdg_plot` object and try to take notice of all of the things that change. 

```{r}
# add theme minimal
sdg_plot +
  XX
```

# 3 - Theme

1. Now use the `theme()` function to change how the legend is displayed. Change the legend position with `legend.position = c(.85, .5)`, remove the legend title using `legend.title = element_blank()`, increase the font size in the legend using `legend.text = element_text(size = 10)`, and set the legend background to white using `legend.background = element_rect(fill = "white", color = "white")`

```{r}
# adjust theme settings
sdg_plot +
  XX +
  theme(XX, XX, XX, XX)
```

# 3 - Colors

1. Add `scale_color_colorblind()` to change the color set used to color the two sets of points belonging to the two SDGs. 

```{r}
# add new color set
sdg_plot +
  XX +
  theme(XX, XX, XX, XX) +
  XX
```

2. Within `scale_color_colorblind()` set `labels = c("No poverty (SDG 1)", "Clean water (SDG 6)")` in order to adjust the legend labels accordingly.  

```{r}
# add new legend labels
sdg_plot +
  XX +
  theme(XX, XX, XX, XX) +
  XX(XX)
```

# 4 - Labs

1. Ok, the plot looks kind of OK by now. What's missing is appropriate annotion. Use `labs()` to add suitable text for the plot title and subtitle, as well as the x and y axis titles.  

```{r}
# add labs
sdg_plot +
  XX +
  theme(XX, XX, XX, XX) +
  XX(XX) +
  labs(title = XX,
       subtitle = XX,
       x = XX,
       y = XX)
```

# 5 - Save plot

1. It's time to save the plot as a `.png`. Before doing this, it makes sense to again save the plot as an object. See code below.  

```{r}
# save as new object
sdg_plot_ready <- sdg_plot +
  XX +
  theme(XX, XX, XX, XX) +
  XX(XX) +
  labs(title = XX,
       subtitle = XX,
       x = XX,
       y = XX)
```

2. Now, use the code below to save your plot into the `3_figures` folder in the project folder. Don't forget to replace the `XX` with your plot object. Also, make sure that you are working in the `SDGDataViz` project and that `3_figures` exists in your working directory (test using `dir()`).   

```{r}
# save plot
ggsave(filename = "3_figures/my_sdg_plot.png", 
       plot = XX, 
       device = "png",
       width = 7, 
       height = 7,
       dpi = 300)
```

3. The settings in the `ggsave` function probably were quite reasonable, but feel free to change arguments such as `width` or `height` to alter the appearance of your `png` plot. 



