---
title: "Wrangling"
output:
editor_options:
  chunk_output_type: console
---


In this practical, you will read in the `sdg_hackathon_data.zip` data set and analyze it descriptively using the `dplyr` package.

**To complete the practical, you will have to replace the `XX` in the code with the correct object or functions and then run the code in the console.** 

# 0 - Preliminaries

1. Run the code below to load the `tidyverse` and `lubridate` packages. If you have not previously installed the `tidyverse` or `lubridate`, first run `install.packages("tidyverse")` and `install.packages("lubridate")`, respectively, and then run the code below. 

```{r }
library(tidyverse)
library(lubridate)
```

2. Use the code below to read in the `sdg_hackathon_data.zip` data set and store it as an object called `project_sdgs`. The messages printed in the console tell you how R has inferred the types of each of the columns. Should be all good.

```{r }
project_sdgs <- read_csv('1_Data/sdg_hackathon_data.zip')
```

3. Take a look at `project_sdgs` by running it in the console. Note how the `tibble` shows the data types for each of the columns. Also note how some cases and variables are abbreviated. 

```{r }
project_sdgs
```

# 1 - Add year

1. Use mutate and `lubridate`'s `year()` function to extract the year from `start_date` variable and store it as a new variable in the data frame called `year`. 

```{r }
# add year variable
project_sdgs %>%
  mutate(year = year(start_date))
```

# 2 - Number of projects per year

1. First, use the `dplyr` functions `select()` and `distinct()` to produce a tibble that only includes the distinct entries of `project_number` and `project_title`. Note how the number of rows displayed at the very top of the print has shrunk to `32,375`. 

```{r}
# distinct projects
project_sdgs %>%
  mutate(year = year(start_date)) %>% 
  select(project_number, project_title) %>% 
  distinct()
```

2. Now, use the `dplyr` functions `summarize()` and `n()` to confirm the number of distinct rows in the data set.  

```{r }
# number of distinct projects
project_sdgs %>%
  mutate(year = year(start_date)) %>% 
  select(project_number, project_title) %>% 
  distinct() %>% 
  summarize(n = n())
```

3. That worked. Now determine the number of projects per year. To do this, first additionally include `year` in the `select` function. Secons, insert a `group_by()` in between the `distinct()` and the `summarize()` functions with `year` as the grouping variable.

```{r }
# number of distinct projects per year
project_sdgs %>%
  mutate(year = year(start_date)) %>% 
  select(project_number, project_title, year) %>% 
  distinct() %>% 
  group_by(year) %>% 
  summarize(n = n())
```

4. You have determined the number of projects for each year, but the print might have only shown the first ten rows. You can tell R to print more rows using the `print()` function. This should show reveal there are not only very few projects that started prior to `2009`, but also that will start after `2021`. For analyses depending on year, it therefore may make sense to eliminate these cases. 
```{r }
# number of distinct projects per year
project_sdgs %>%
  mutate(year = year(start_date)) %>% 
  select(project_number, project_title, year) %>% 
  distinct() %>% 
  group_by(year) %>% 
  summarize(n = n()) %>% 
  print(n = 20)
```

# 3 - Filtering projects

1. Use `dplyr`'s `filter` function to retain only those projects that have `year` larger `2008` and smaller `2022`. 

```{r }
# number of distinct projects per year
project_sdgs %>%
  mutate(year = year(start_date)) %>% 
  select(project_number, project_title, year) %>% 
  distinct() %>% 
  group_by(year) %>% 
  summarize(n = n())  %>%
  filter(year > 2008, year < 2022)
```

# 4 - Proportion of SDGs per system

1. Can you figure out how to use the code below to determine the proportion of SDGs per system?

```{r }
# proportion of SDGs per system
project_sdgs %>%
  group_by(sdg, system) %>% 
  summarize(n = n()) %>% 
  group_by(system) %>% 
  mutate(proportion = n / n()) %>% 
  print(n = 100)
```





